### 2018-05-07 Writing/disparity_review.tex
tests section

\subsection{Tests}
Similarly to all the points covered above, which statistical test to apply solely depends on which question is tackled with disparity analysis (i.e. which hypothesis to test).
On the other hand, some tests (such as permanovas [CITE]) are actually \textit{not} taking into account the disparity measured but are simply based on the distribution of the taxa into the space (or simply their pairwise distances in the case of PERMANOVAS).

Furthermore, bootstraps are often used in morphological disparity analysis as a mean to get a distribution from a single disparity value [CITE].
While this method is valid in itself, it comes with several caveats.
First, it is important to consider that bootstrapping the disparity values creates a distribution of them (e.g. a distribution of sum of variances from a bootstrapped morphospace).
Therefore, the error associated with the bootstrap is reflecting the ``stability'' of the metric, not the extremes of the distribution of the data (\textit{c.f.} when bootstrapping a disparity distribution; e.g. the distribution of all the distances from centroids from a bootstrapped morphospace).
In the first and most common case, the amount of bootstraps artificially increases the type I error.
Furthermore, when comparing two groups, the comparison is one of the disparity of two groups using their ``stability'' as variance.

Best practices:
\begin{itemize}
\item make sure your metrics are capturing what you are testing
\item check out the test assumptions
\item Know if and how your dataset is violating assumptions.
\end{itemize}

\subsubsection{Models}
What are our null hypotheses?
Joe’s simulations’ stuff - phylogeny can be used to build a null.One basic null is that disparity increases with diversity; how do we diverge from that (show these morpho/genome distance divergence/convergence).
We can have a null model that is based on simple Brownian increase in disparity (gradual). 
Null: is the null increase in disparity dependent on a random walk and on a tree shape?
Maybe we don’t want to think about a null model but rather think about good models to fit. 
We haven’t explored much about fitting models like explaining holes, effects of extinction etc...
Dealing with uncertainty? Bootstrap Bayesian G matrices \& P matrices!

@@

\section{Where do we go from there?}
4 - This a list of problems for the future/moving forward
    Simulations 
    Form/function
Isospaces, glottospaces, acoustospace, ecospace, life-history-spaces, etc. All just a multidimensional analysis? Needs for terms?
    
``We asked whether we could but not whether we should'' (Jurassic Park)

>>>>

\subsection{Tests}
Similarly to all the points covered above, the statistical test to apply solely depends on which question is tackled with disparity analysis (i.e. which hypothesis to test).
Considering the question, the data, the ordination method used (or the absence thereof) and the metric used, the statistical test should be based on a clearly state hypothesis.
This might seem obvious but we believe doing so greatly enhances the framing of the analysis and its comprehension by future readers.

For example, one commonly used test in morphological analysis, is the non parametric permutation anovas [NPANOVA or PERMANOVA; CITE Anderson + examples].
This test and its implementations is perfectly valid to test whether two groups have a different variance/covariance but is based on the pairwise distances of the group elements.
This test will ignore any eventual ordination (if the data was ordinated) and is not based on a uni-dimensional disparity metric.
In other words, if one observes no overlap between two groups on the two first axis of a PCA, the hypothesis H0 that the two groups overlap in the morphospace is not directly tested by a PERMANOVA which tests rather whether the two groups share the same variance/covariance based in a ``distance-space''.
Therefore, explicitly stating the hypothesis might help understanding which test to apply depending on the disparity analysis question.

Another caveat for applying statistical tests on disparity data is on which data to actually apply the test.
In fact, in morphological disparity analysis (especially for palaebiological questions), data is often bootstrapped.
This has two useful advantages: first, when the disparity metric is unidimensional (e.g. the sum of variances or the average squared pairwise distance), bootstrapping the data will generate a distribution of this unidimensional metric that can then be analysed using the vast statistical toolkit to compare distributions; second, when the data is scarce (e.g. for palaeontological data), bootstrapping the data will allow to introduce statistical variance and will render the test less sensitive to outliers.
However, bootstrapping the data also comes with an important caveat: the data analysed is pseudo-replicate and thus non-independant.
This can violate most of the parametric statistical tests assumptions.
Furthermore, it is important to remember that the number of bootstrap pseudo-replicates will inevitably increase the type I error.

\section{Future directions}

@@

\subsubsection{Disparity models}
What are our null hypotheses?
Joe’s simulations’ stuff - phylogeny can be used to build a null.One basic null is that disparity increases with diversity; how do we diverge from that (show these morpho/genome distance divergence/convergence).
We can have a null model that is based on simple Brownian increase in disparity (gradual). 
Null: is the null increase in disparity dependent on a random walk and on a tree shape?
Maybe we don’t want to think about a null model but rather think about good models to fit. 
We haven’t explored much about fitting models like explaining holes, effects of extinction etc...
Dealing with uncertainty? Bootstrap Bayesian G matrices \& P matrices!

\subsection{Where do we go from there?}
4 - This a list of problems for the future/moving forward
    Simulations 
    Form/function
Isospaces, glottospaces, acoustospace, ecospace, life-history-spaces, etc. All just a multidimensional analysis? Needs for terms?

%Concluding words:
Although disparity analysis have been rendered easy to perform from the user end (CITE packages tools), it is crucial to remember that they are first and foremost multidimensional analysis; and multidimensional analysis are complex.
In Jurassic Park, Dr Ian Malcolm summarises this problem in an elegant way: ``[We] scientists were so preoccupied with whether or not [we] could that [we] didn't stop to think if [we] should.''
We believe that morphological analysis in general will greatly benefit in emphasising the methodological decisions (\textit{whether we should}) rather than simply using disparity analysis because \textit{we can}.
